---
title: "Homework 3"
author: "Ashley Cooper, Chloe Swick, Eleanor Thomas"
date: "Due 5/19/23"
output:
  html_document:
    code_folding: hide
    theme: flatly
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE, results = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results = FALSE)

library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(cowplot)
```

### Question 1

#### Run linear regressions (with an intercept) to estimate the demand curves for DAC and non-DAC consumers. What are 2-3 reasons you can think that may explain the differences in demand you estimate?

```{r, results = TRUE}
ev_clean <- read_csv(here("HW3_data.csv")) %>% 
  janitor::clean_names() %>% 
  rename("price" = "price_dollars",
         "dac" = "q_dac",
         "non_dac" = "q_non_dac")

non_dac_lm <- lm(price ~ non_dac, data = ev_clean)
non_dac_coeff <- non_dac_lm$coefficients
###   (Intercept)       non_dac 
### 158034.280239     -2.730926 

dac_lm <- lm(price ~ dac, data = ev_clean)
dac_coeff <- dac_lm$coefficients
###  (Intercept)           dac 
### 116854.835741     -6.611437 

ev_pivot <- ev_clean %>% 
  pivot_longer(cols = dac:non_dac, names_to = "community", values_to = "quantity")

demand_plot <- ggplot(data = ev_pivot) +
  geom_abline(slope = -2.730926, intercept = 158034.280239) +
  geom_abline(slope = -6.611437, intercept = 116854.835741) +
  geom_point(aes(x = quantity, y = price, color = community)) +
  scale_color_viridis_d(labels = c("DAC", "Non-DAC")) +
  labs(x = "Number of EVs", y = "Maginal Cost ($/car)", color = "Community Type") +
  ggtitle("Demand for Electric Vehicles (EVs)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
  
demand_plot
```

Price = 116854.835741 - 6.611437(QDAC)

6.611437(QDAC) = 116854.835741 - Price



[*QDAC = (116854.835741 - Price)/6.611437*]{.underline}

Price = 158034.280239 - 2.730926(QNonDAC)

2.730926(QNonDAC) = 158034.280239 - Price

[*QNonDAC = (158034.280239 - Price)/2.730926*]{.underline}

### Question 2

#### One kilogram (kg) of lithium is estimated to cause \$300 dollars worth of damage in health costs to local communities living near mines. Each EV requires a lithium ion battery containing 63kg of lithium. Assuming there are no other externalities caused by EV production, what is the marginal externality cost per EV produced?

[*MEC = \$300/kg \* 63kg/car MEC = \$18,900*]{.underline}

### Question 3

#### What is the aggregate annual demand curve for EVs? What is the supply curve for EVs?

```{r}
q_dac <- function(price){
    quantity = (116854.835741 - price)/6.611437
    return(quantity)
}

q_non_dac <- function(price){
    quantity = (158034.280239 - price)/2.730926
    return(quantity)
}

q_agg <- function(price){
  if(price > 116854.835741) {
    quantity = (158034.280239 - price)/2.730926
    return(quantity)
    }
if(price <= 116854.835741) {
  quantity = q_dac(price) + q_non_dac(price)
  return(quantity)
}

}


# Qagg = (116854.835741 - price)/6.611437) + (158034.280239 - price)/2.730926
#         17674.65 - price/6.611437 + 57868.39 - price/2.730926
#         75543.04 - -0.5174292 * price



### The current EV price (without any intervention) is $50,000 per vehicle.
### To get the slope of the Supply Curve, P = a + bQ
### b = (P2 - P1)/(Q2 - Q1)
### Goes through the intercept, so (Q1, P1) = (0,0) ; P2 = $50,000 (current EV price) and Q1 = Qagg($50,000)

# q_agg(50000)
# 49671.58

supp_slope <- (50000)/q_agg(50000)
# supp_slope
### 1.006612
```

Aggregate Annual Demand Equation
If price > 116854.835741:
    Qagg = (158034.280239 - Price)/2.730926

If price <= 116854.835741):
    Qagg = 75543.04 - -0.5174292(Price)

Supply Equation
Price = 1.006612(Qagg)

#### What is the "benefit" to producers under the status quo? What is the environmental cost under the status quo?

```{r}
cs <- 1/2*(116854.835741 - 50000)*q_dac(50000) + 1/2*(158034.280239 - 50000)*q_non_dac(50000)
# cs
# 2474913243

ps <- 1/2*50000*49671.58
# ps
# 1241789500
```

Benefit to Consumers = Consumer Surplus = $2,474,913,243

Benefit to Producers = Producer Surplus = $1,241,789,500

#### What is the environmental cost under the status quo?

```{r}
### Environmental cost = MEC ($18,900) * Qagg(50,000)
env_cost <- 18900 * 49671.58
# env_cost
# 938792862
```
Environmental Costs = $938,792,862

### Question 4
#### How is the current consumer benefit divided between DAC and non-DAC consumers?
```{r}
cs_dac <- 1/2*(116854.835741 - 50000)*q_dac(50000)
#cs_dac -- $338,017,973
cs_non_dac <- 1/2*(158034.280239 - 50000)*q_non_dac(50000)
#cs_non_dac -- $2,136,895,270

consumer_benefit_dac <- cs_dac/cs
consumer_benefit_non_dac <- cs_non_dac/cs

# consumer_benefit_dac
# 0.1365777
# consumer_benefit_non_dac
# 0.8634223
```

Currently disadvantaged communities get ~13.66% of the total consumer benefits from EVs, while non-disadvantaged communities enjoy ~86.34% of the total consumer benefits.

### Question 5
#### Derive the optimal EV tax (in $ per vehicle) to internalize the lithium mining externality.

Optimal Tax = MEC at Q*

Because MEC is constant, optimal tax = $18,900/EV

#### Noting that recent research has shown that DACs are far more likely to contain mining activity, assume that the mining externality is borne entirely by the DAC group. What would be the effects of this tax on (a) The amount of EVs produced and consumed

```{r}
q_agg(50000)
# 49671.58

q_agg(68900)
# 39892.16

q_agg(50000) - q_agg(68900)
# 9779.412
```
The quantity of EVs would decrease by **~ 9,780** cars, from 49,672 EVs to **39,892 EVs**.

#### (b) The price of EVs

The price of each EV would increase by **$18,900** from $50,000 to **$68,900**.

#### (c) Overall welfare of non-DAC consumers
```{r}
cs_non_dac_tax <- 1/2*(158034.280239 - 68900)*q_non_dac(68900)
# 1,454,620,139
```

Benefit to non-DAC Consumers w/tax = $1,454,620,139
This is a decrease of **$682,275,131** from a non taxed CS of $2,136,895,270 


#### (d) Overall welfare of DAC consumers
```{r}
cs_dac_tax <- 1/2*(116854.835741 - 68900)*q_dac(68900)
cs_dac_tax
# 173,915,767
```

Benefit to DAC Consumers w/tax = **$173,915,767**
This is a decrease of **$164,102,206** from the non taxed CS of $338,017,973 


#### (e) Benefit to EV producers
```{r}
ps <- 1/2*50000*49671.58
# ps
# 1241789500
ps_tax <- 1/2*68900*39892.16
#1,374,284,912
```

Benefit to Producers w/tax = **$1,374,284,912**
This is a decrease of **$ 132,495,412**from the non taxed PS of $1,241,789,500 


#### (f) Total environmental damage
```{r}
### Environmental cost = MEC ($18,900) * Qagg(50,000)
#q_agg(68900)
# 39892.16
env_cost_tax <- 18900 * 39892.16

# env_cost no tax
# 938792862
#env_cost_tax
#753,961,824
```

Environmental Damage w/tax = **$753,961,824**
This is a decrease of **$184,831,038** from the non taxed PS of $938,792,862 


#### (g) Total tax revenue generated
<br>
Total tax revenue generated is the same as the total environmental damage, $184,831,038 because the optimal tax was evaluated to cover all of the environmental damage.

6. Now, assume that all revenue from the EV tax will be redistributed to the consumers in proportion to their pre-tax consumption. For example, if 80% of the EVs were consumed by non-DAC consumers, then they get 80% of the tax revenue. Additionally, consider that emerging scientific evidence suggests the true environmental damages from lithium mining may be much higher than $300. For a range of values of external costs per kg of lithium ($350, $400, $450, and $500 per kg), calculate the effects of an EV tax on:1

