---
title: "Homework 3"
author: "Ashley Cooper"
date: "Due 5/19/23"
output:
  html_document:
    code_folding: hide
    theme: flatly
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE, results = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results = FALSE)

library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(cowplot)
```

### Question 1

#### Run linear regressions (with an intercept) to estimate the demand curves for DAC and non-DAC consumers. What are 2-3 reasons you can think that may explain the differences in demand you estimate?

```{r, results = TRUE}
ev_clean <- read_csv(here("HW3_data.csv")) %>% 
  janitor::clean_names() %>% 
  rename("price" = "price_dollars",
         "dac" = "q_dac",
         "non_dac" = "q_non_dac")

non_dac_lm <- lm(price ~ non_dac, data = ev_clean)
non_dac_coeff <- non_dac_lm$coefficients
###   (Intercept)       non_dac 
### 158034.280239     -2.730926 

dac_lm <- lm(price ~ dac, data = ev_clean)
dac_coeff <- dac_lm$coefficients
###  (Intercept)           dac 
### 116854.835741     -6.611437 

ev_pivot <- ev_clean %>% 
  pivot_longer(cols = dac:non_dac, names_to = "community", values_to = "quantity")

demand_plot <- ggplot(data = ev_pivot) +
  geom_abline(slope = -2.730926, intercept = 158034.280239) +
  geom_abline(slope = -6.611437, intercept = 116854.835741) +
  geom_point(aes(x = quantity, y = price, color = community)) +
  scale_color_viridis_d(labels = c("DAC", "Non-DAC")) +
  labs(x = "Number of EVs", y = "Maginal Cost ($/car)", color = "Community Type") +
  ggtitle("Demand for Electric Vehicles (EVs)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
  
demand_plot
```

Price = 116854.835741 - 6.611437(QDAC)

6.611437(QDAC) = 116854.835741 - Price



[*QDAC = (116854.835741 - Price)/6.611437*]{.underline}

Price = 158034.280239 - 2.730926(QNonDAC)

2.730926(QNonDAC) = 158034.280239 - Price

[*QNonDAC = (158034.280239 - Price)/2.730926*]{.underline}

### Question 2

#### One kilogram (kg) of lithium is estimated to cause \$300 dollars worth of damage in health costs to local communities living near mines. Each EV requires a lithium ion battery containing 63kg of lithium. Assuming there are no other externalities caused by EV production, what is the marginal externality cost per EV produced?

[*MEC = \$300/kg \* 63kg/car MEC = \$18,900*]{.underline}

### Question 3

#### What is the aggregate annual demand curve for EVs? What is the supply curve for EVs?

```{r}
q_dac <- function(price){
    quantity = (116854.835741 - price)/6.611437
    return(quantity)
}

q_non_dac <- function(price){
    quantity = (158034.280239 - price)/2.730926
    return(quantity)
}

q_agg <- function(price){
  if(price > 116854.835741) {
    quantity = (158034.280239 - price)/2.730926
    return(quantity)
    }
if(price <= 116854.835741) {
  quantity = q_dac(price) + q_non_dac(price)
  return(quantity)
}

}


# Qagg = (116854.835741 - price)/6.611437) + (158034.280239 - price)/2.730926
#         17674.65 - price/6.611437 + 57868.39 - price/2.730926
#         75543.04 - -0.5174292 * price



### The current EV price (without any intervention) is $50,000 per vehicle.
### To get the slope of the Supply Curve, P = a + bQ
### b = (P2 - P1)/(Q2 - Q1)
### Goes through the intercept, so (Q1, P1) = (0,0) ; P2 = $50,000 (current EV price) and Q1 = Qagg($50,000)

# q_agg(50000)
# 49671.58

supp_slope <- (50000)/q_agg(50000)
# supp_slope
### 1.006612
```

Aggregate Annual Demand Equation
If price > 116854.835741:
    Qagg = (158034.280239 - Price)/2.730926

If price <= 116854.835741):
    Qagg = 75543.04 - -0.5174292(Price)

Supply Equation
Price = 1.006612(Qagg)

#### What is the "benefit" to producers under the status quo? What is the environmental cost under the status quo?

```{r}
cs <- 1/2*(116854.835741 - 50000)*q_dac(50000) + 1/2*(158034.280239 - 50000)*q_non_dac(50000)
# cs
# 2474913243

ps <- 1/2*50000*49671.58
# ps
# 1241789500
```

Benefit to Consumers = Consumer Surplus = $2,474,913,243

Benefit to Producers = Producer Surplus = $1,241,789,500

#### What is the environmental cost under the status quo?

```{r}
### Environmental cost = MEC ($18,900) * Qagg(50,000)
env_cost <- 18900 * 49671.58
# env_cost
# 938792862
```
Environmental Costs = $938,792,862

### Question 4
#### How is the current consumer benefit divided between DAC and non-DAC consumers?
```{r}
cs_dac <- 1/2*(116854.835741 - 50000)*q_dac(50000)
cs_non_dac <- 1/2*(158034.280239 - 50000)*q_non_dac(50000)

consumer_benefit_dac <- cs_dac/cs
consumer_benefit_non_dac <- cs_non_dac/cs

# consumer_benefit_dac
# 0.1365777
# consumer_benefit_non_dac
# 0.8634223
```

Currently disadvantaged communities get ~13.66% of the total consumer benefits from EVs, while non-disadvantaged communities enjoy ~86.34% of the total consumer benefits.

### Question 5
#### Derive the optimal EV tax (in $ per vehicle) to internalize the lithium mining externality. Noting that recent research has shown that DACs are far more likely to contain mining activity, assume that the mining externality is borne entirely by the DAC group. What would be the effects of this tax on (a) The amount of EVs produced and consumed

```{r}
### Optimal Tax = MEC at Q*
### Because MEC is constant, optimal tax = $18,900/EV
```


#### (b) The price of EVs

The price of each EV would increase by $18,900.

#### (c) Overall welfare of non-DAC consumers

#### (d) Overall welfare of DAC consumers

#### (e) EV producers

#### (f) Total environmental damage

#### (g) Total tax revenue generated
